---
title: 'Reproducible Research Project: Technical and Humanities Students’ Perspectives on the Development and Sustainability of Artificial Intelligence (AI)'
author: 'Szymon Karkoszka, Antoni Piotrowski'
format: 
  html:
    code-fold: true
    toc: true
    toc-depth: 2
    toc-expand: 1
    toc-title: Contents
    toc-location: body
    smooth-scroll: true
    theme:
      light: flatly
      dark: darkly
title-block-banner: true
output:
  html_document: default
  pdf_document: default
---


# 1. Why it is important to reproduce the paper — Problem Description and Topic Importance

The perception of students on AI in education is a crucial topic to analyze from an economic standpoint. The incorporation of AI in the education sector has been increasing in recent years, with the emergence of smart classrooms and personalized learning platforms. The potential benefits of AI in education include improved learning outcomes, increased accessibility, and efficiency gains in educational delivery.

Understanding the perceptions of students towards AI in education is vital to inform policies and investment decisions in the education sector. The analysis of student perceptions can provide insights into the potential barriers and facilitators to AI adoption in education, as well as inform strategies for the development of AI technologies that are sensitive to student needs and concerns.

This paper aims to analyze students’ perception on the AI in Education. Therefore, we decide to examine the differences between the studies.

## Required Packages

Click on the arrow to the right to see the required R packages used in this analysis.

<details>
<summary>Load Required Libraries</summary>

```{r}
# Loading necessary libraries for analysis

library(tidyverse)        # Data manipulation and visualization
library(RColorBrewer)     # Color palettes
library(reshape2)         # Data reshaping
library(ggplot2)          # Creating graphics
library(kableExtra)       # Enhancing tables
library(scales)           # Scale functions for visualization
library(dplyr)            # Data manipulation
library(tidyr)            # Data tidying
library(knitr)            # Automatic reporting
library(kableExtra)       # Enhancing tables
library(car)              # Advanced regression models
library(rstudioapi)       # Data manipulation and visualization
library(plotly)           # Data manipulation and visualization
```     


# 2. Data

In terms of the comparison, the data we found are excellent as they both compare students and are both from Romania. Unfortunately, our dataset is *10 times smaller* but much more informative. Our data are also much more up to date, and we can say that we compare data between two "AI Waves" — one from 2017 and another from 2022.

The dataset utilized in this study encompasses the responses obtained from a survey administered to undergraduate students in their 2nd and 3rd years of study at the Faculty of Cybernetics, Statistics, and Economic Informatics. The survey was conducted online and disseminated through social media groups. Its objective was to acquire insights into students' perceptions regarding the role of artificial intelligence (AI) in the domain of education.

This dataset comprises a total of 96 observations, each representing a unique case or participant. Additionally, the dataset contains 36 variables, which encompass the different characteristics or attributes under investigation.

## 2.1 Loading the Data

Click on the arrow to the right to view the code used for loading the data.

<details>   
<summary>Load Dataset from Local Storage</summary>
```{r}
current_path <- rstudioapi::getActiveDocumentContext()$path
current_path <- path.expand(current_path)
current_path <- dirname(current_path)
to_csv <- file.path(current_path, "Data", "Survey_AI.csv")
to_csv
```

```{r}
data <- read.csv(to_csv)
```
</details>



## 2.2 List of Variables

This section provides a description of the variables used in the study. The main dependent variable is concerned with the perceived utility of AI in the educational process.

### Dependent Variable

- **Q7.Utility_grade**: On a scale of 1 to 10, how useful do you think AI would be in the educational process? 
  - 1: Not useful at all
  - 10: Extremely useful

### Independent Variables

#### Identification
- **ID**: Student ID

#### Knowledge and Sources about AI
- **Q1.AI_knowledge**: Level of knowledge about artificial intelligence on a scale of 1 to 10
  - 1: Not informed at all
  - 10: Extremely informed
- **Q2.AI_sources**: Sources used to learn about artificial intelligence:
  - **Q2.1.Internet**: Whether Internet was used as a source (1 = Yes, 0 = No)
  - **Q2.2.Books.Papers**: Whether Books/Scientific papers were used (1 = Yes, 0 = No)
  - **Q2.3.Social_media**: Whether social media was used (1 = Yes, 0 = No)
  - **Q2.4.Discussions**: Whether discussions with family/friends were used (1 = Yes, 0 = No)
  - **Q2.5.NotInformed**: Indicator if "I don't inform myself about AI" was chosen (1 = Yes, 0 = No)

#### Opinions on AI
- **Q3.1.AI_dehumanization**: AI encourages dehumanization (1-5 scale)
- **Q3.2.Job_replacement**: Robots will replace people at work (1-5 scale)
- **Q3.3.Problem_solving**: AI helps to solve many problems in society (1-5 scale)
- **Q3.4.AI_rulling_society**: AI will rule society (1-5 scale)
- **Q4.1.AI_costly**: Machinery using AI is very expensive to build and maintain (1-5 scale)
- **Q4.2.Economic_crisis**: AI will lead to a global economic crisis (1-5 scale)
- **Q4.3.Economic_growth**: AI will help global economic growth (1-5 scale)
- **Q4.4.Job_loss**: AI leads to job losses (1-5 scale)

#### Personal Feelings and Impact Domains
- **Q5.Feelings**: Feelings associated with AI
  - 1: Curiosity
  - 2: Fear
  - 3: Indifference
  - 4: Trust
- **Q6.Domains**: Areas where AI would have a big impact
  - **Q6.1.Education**: (1 = Yes, 0 = No)
  - **Q6.2.Medicine**: (1 = Yes, 0 = No)
  - **Q6.3.Agriculture**: (1 = Yes, 0 = No)
  - **Q6.4.Constructions**: (1 = Yes, 0 = No)
  - **Q6.5.Marketing**: (1 = Yes, 0 = No)
  - **Q6.6.Administration**: (1 = Yes, 0 = No)
  - **Q6.7.Art**: (1 = Yes, 0 = No)

#### Educational Process
- **Q8.Advantage_teaching**: Main advantage of AI in the teaching process (1-4 scale)
- **Q9.Advantage_learning**: Level of advantage of AI in the learning process (1-4 scale)
- **Q10.Advantage_evaluation**: Level of advantage of AI in the evaluation process (1-4 scale)
- **Q11.Disadvantage_educational_process**: Level of disadvantage of AI in the educational process (1-4 scale)

#### Demographics and Academic Performance
- **Q12.Gender**: Gender of the student (1 = Male, 2 = Female)
- **Q13.Year_of_study**: Year of study of the student (1-Year 2, 2-Year 3)
- **Q14.Major**: Major or field of study of the student
  - 1: Economic Cybernetics
  - 2: Statistics and Economic Forecasting
  - 3: Economic Informatics
- **Q15.Passed_exams**: Indicator if the student passed the exams (1 = Yes, 0 = No)
- **Q16.GPA**: Student's GPA for the last year of study on a scale from 1 to 10

# 4. Further Data Analysis and Visualization

Continuing with our analysis, we now focus on transforming the data and examining it for completeness, followed by a comparative study with previous findings.

## Examining Data Structure

To understand the variables and their types better, check the data structure by expanding the section below.

<details>
<summary>View Data Structure</summary>

```{r}
# Show the structure of the data to verify variable types and completeness
str(data)
```
</details>
### Cleaning Data
Removing Unnecessary Variables
Since some character variables are only annotations, they will be removed. See the details of this process below.

<details>
<summary>Remove Character Variables</summary>
```{r}
# Removing character variables that serve as comments for certain responses
data <- subset(data, select = -c(Q2.AI_sources, Q6.Domains, ID))
```
</details>
### Converting Variables to Numeric
To facilitate our analysis, all variables need to be numeric. This conversion can be reviewed below.

<details>
<summary>Convert All Variables to Numeric</summary>
```{r}
# Converting all variables in the dataset to numeric for consistent analysis
data <- mutate_all(data, as.numeric)
```
</details>
### Checking for Missing Values
Before proceeding further, it's essential to ensure there are no missing values in the dataset.

<details>
<summary>Check for Missing Values</summary>
```{r}
# Check and print the sum of missing values across all columns
print(missing_values <- colSums(is.na(data)))
```
</details>


# 5. Visual Analysis of Student Knowledge

With no missing values confirmed, we proceed to a visual representation of student knowledge on AI.

The provided plot shows the distribution of AI knowledge ratings among students. It is observed that:

- The ratings of 7.5 and 5.0 are most frequent, indicating a moderate to good understanding of AI by the majority.
- Extremely low (2.5) and high (10.0) ratings are less frequent, reflecting that fewer students consider their understanding of AI to be very poor or excellent.
- The distribution centers around mid to slightly higher knowledge ratings, with fewer students at the extremes.

This visualization is crucial for understanding the spread of AI knowledge levels across the surveyed student population and helps in identifying areas where additional education or clarification might be needed.

<details>
<summary>Click here to view the code used for generating the plot</summary>

```{r}
custom_colors <- brewer.pal(10, "Spectral")

fig1 <- ggplot(data, aes(x = Q1.AI_knowledge)) +
  geom_bar(aes(fill = factor(Q1.AI_knowledge)), color = "black") +
  scale_fill_manual(values = custom_colors) +
  xlab("AI knowledge rating") +
  ylab("Number of responses") +
  ggtitle("A countplot of AI knowledge ratings") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 12),
        axis.text = element_text(size = 10),
        legend.title = element_blank(),
        legend.position = "none")


```
</details>
```{r}
print(fig1)
```


# 6. Analysis of Information Sources

We analyzed the sources students use to inform themselves about artificial intelligence, which is crucial for understanding their exposure and potentially the depth of their knowledge.

The bar plot below illustrates the proportion of students using each source:

- **Internet**: The most utilized source, used by nearly 40% of the students.
- **Social Media**: The second most popular source, with about 25% usage.
- **Books/Papers**: Follows closely behind social media, used by roughly 20% of the students.
- **Discussions**: Less common, with around 15% of students engaging in discussions about AI.
- **Not Informed**: A small proportion, approximately 5%, do not seek information about AI.

This distribution suggests a heavy reliance on digital sources, with traditional sources like books and papers still playing a significant role.
<details>
<summary>Click here to view the code used for generating the plot</summary>
```{r}
source_names <- c("Internet", "Books/Papers", "Social Media", "Discussions", "Not Informed")
source_counts <- sapply(data[, c("Q2.1.Internet", "Q2.2.Books.Papers", "Q2.3.Social_media", "Q2.4.Discussions", "Q2.5.NotInformed")], function(col) sum(col == 1))

source_data <- data.frame(Source = source_names, Count = source_counts)
source_data$Proportion <- source_data$Count / sum(source_data$Count)

custom_colors <- brewer.pal(5, "Spectral")
fig2 <- ggplot(source_data, aes(x = reorder(Source, -Proportion), y = Proportion, fill = Source)) +
  geom_bar(stat = "identity", color = "black") +
  labs(x = "Source", y = "Proportion", fill = "Source") +
  ggtitle("Proportion of Students Using Each Source") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 12),
        axis.text = element_text(size = 10),
        legend.title = element_text(size = 12),
        legend.position = "right") +
  scale_fill_manual(values = custom_colors)

```
</details>
```{r}
print(fig2)
```

# 7. Correlation Analysis Among Information Sources

This section explores the relationships between different sources students use to inform themselves about artificial intelligence. A correlation heatmap provides a visual representation of these relationships.

The heatmap below shows how each source of information correlates with others. Positive correlations (shown in blue) suggest that students who use one source are likely to use another, whereas negative correlations (shown in red) suggest that students who use one source are less likely to use another.



<details>
<summary>Click here to view the code used for generating the correlation heatmap</summary>

```{r}
selected_columns <- c("Q2.1.Internet", "Q2.2.Books.Papers", "Q2.3.Social_media", "Q2.4.Discussions", "Q2.5.NotInformed")
selected_data <- data[,selected_columns]

# Calculate the correlation matrix and round the values
correlation_matrix <- round(cor(selected_data), 2)
correlation_data <- melt(correlation_matrix)

# Genrate heatMap
corelationHeatMap <- ggplot(data = correlation_data, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "red", mid = "white", high = "blue", midpoint = 0) +
  theme_minimal() +
  labs(title = "Correlation Heatmap", x = "", y = "") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```
</details>

```{r}
print(corelationHeatMap)
```


# 8. Relation of Student Knowledge about AI to Their Views on Various AI-related Statements

This section explores how students' knowledge about AI correlates with their opinions on various AI-related statements. The analysis aims to discern patterns or trends in agreement levels across different knowledge segments.

## 8.1 Analysis of Mean Knowledge Scores by Information Source Usage

We first analyze the mean AI knowledge scores for students based on whether they use specific information sources. This provides insights into whether accessing particular types of information correlates with higher AI knowledge.

<details>
<summary>Click here to view the code used for this analysis</summary>

```{r}
# Define the columns vector
columns <- c("Q2.1.Internet", "Q2.2.Books.Papers", "Q2.3.Social_media", "Q2.4.Discussions", "Q2.5.NotInformed")

# Create an empty list to store the results
results <- list()

# Iterate over the columns
for (i in columns) {
  x1 <- mean(data[data[, i] == 0, "Q1.AI_knowledge"], na.rm = TRUE)
  x2 <- mean(data[data[, i] == 1, "Q1.AI_knowledge"], na.rm = TRUE)
  change <- round(x2 - x1, 2)
  results[[i]] <- c(x1, x2, change)
}

# Create a data frame from the results list
data2 <- data.frame(t(as.data.frame(results)))
colnames(data2) <- c("No - MeanQ1", "Yes - MeanQ1", "Absolute Change")
row.names(data2) <- columns
```

## 8.2 Visualization of Agreement Levels on AI Statements

The violin plot below visualizes students' levels of agreement with various statements about AI, ranging from its impact on society and work to its economic implications.


<details>
<summary>Click here to view the code used for generating the violin plot</summary>

```{r, echo=TRUE,warning=FALSE,message=FALSE,error=FALSE}
# Define the column names and labels
columns2 <- c('Q3.1.AI_dehumanization', 'Q3.2.Job_replacement', 'Q3.3.Problem_solving', 'Q3.4.AI_rulling_society',
                    'Q4.1.AI_costly', 'Q4.2.Economic_crisis', 'Q4.3.Economic_growth', 'Q4.4.Job_loss')
labels <- c('AI encourages dehumanization', 'Robots will replace people at work', 'AI helps to solve many problems in society',
            'AI will rule society', 'Machinery using AI is very expensive',
            'AI will lead to a global economic crisis', 'AI will help global economic growth', 'AI leads to job losses')

# Reshape the data for plotting
data_subset_long <- melt(data)
names(data_subset_long) <- c("variable", "value")
data_subset_long$variable <- factor(data_subset_long$variable, levels = columns2, labels = labels)

# Create a color palette
my_palette <- brewer.pal(8, "Set2")

# Create a violin plot
p <- ggplot(data_subset_long, aes(x = variable, y = value, fill = variable)) +
  geom_violin(trim = FALSE, color = "black", alpha = 0.8) +
  scale_fill_manual(values = my_palette) +
  labs(title = "Violin Plot of Questions 3 & 4 Statements versus Responses",
       x = "Statement", y = "Agreement Level") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
        plot.title = element_text(hjust = 0.5, size = 14),
        legend.position = "none",
        panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank())
```
</details>
```{r}
# Print the plot
print(p)
```



# 9. Analysis of Students' Feelings About the Development of AI

In this section, we explore how students feel about the development of AI, a key factor in understanding their emotional and psychological engagement with this technology.

The pie chart below illustrates the distribution of feelings among students:

- **Curiosity** is the most dominant feeling, representing 68.1% of the responses, indicating a high level of interest and positive engagement with AI.
- **Fear** comes next at 13.2%, suggesting a significant minority of students harbor concerns about AI's implications.
- **Indifference** and **Trust** represent smaller proportions at 11% and 7.7% respectively, showing that fewer students feel neutral towards AI or trust it implicitly.

<details>
<summary>Click here to view the code used for generating the pie chart</summary>

```{r}
# Extract the desired values and labels from the dataframe
feelings <- c('Curiosity', 'Fear', 'Indifference', 'Trust')
# Assuming that data$Q5.Feelings is your vector of feelings
str(data$Q5.Feelings)

print(data$Q5.Feelings)
counts <- table(data$Q5.Feelings)

# Define custom colors for the pie slices
custom_colors <- c('#F8766D', '#00BA38', '#619CFF', '#FFB028')

# Calculate percentages
percentages <- round(counts / sum(counts) * 100, 1)

# Labels with percentages
labels <- paste(percentages, "%\n", feelings, sep="")
```
</details>

```{r}
# Create the pie chart with enhanced appearance
pie(counts, labels = labels, col = custom_colors, border = 'white', 
    init.angle = 90, clockwise = TRUE, main = "Students' Feelings about the Development of AI")

# Add a legend
legend("right", legend = feelings, fill = custom_colors,
       bty = "n", cex = 0.8)
```


# 10. Comparison of Students' Perception of AI

In this section, we compare our dataset's results concerning students' perceptions of AI with another study visualized in the pie chart above. This comparison will help identify any shifts in sentiment or notable differences in how students perceive AI.

## Compared Paper Study Results

Below is a summary of our study results for comparison:

- **Confused**: 26.6%
- **Indifferent**: 10.3%
- **Optimistic**: 28.2%
- **Enthusiastic**: 11.2%
- **Concerned**: 23.7%

The comparison highlights a higher optimism and enthusiasm among our study participants, with lower levels of confusion and concern compared to the mentioned study.
<details>
<summary>Click here to view the code used for generating the pie chart</summary>

```{r}
# Load necessary libraries
library(plotly)

# Data from our study
categories <- c("confused", "indifferent", "optimistic", "enthusiastic", "concerned")
percentages <- c(26.6, 10.3, 28.2, 11.2, 23.7)
values <- percentages / 100 * sum(percentages)  # Assuming total sum of percentages as a whole

# Create a dataframe
data_plot <- data.frame(categories, values)

# Create the pie chart for our study
fig <- plot_ly(data_plot, labels = ~categories, values = ~values, type = 'pie',
               textinfo = 'label+percent',
               insidetextorientation = 'radial',
               marker = list(colors = c('#FF7F0E', '#1F77B4', '#FFBB78', '#AEC7E8', '#2CA02C')))

# Enhance the layout
fig_from_paper <- fig %>% layout(title = 'Our Study: Students Perception of AI',
                      showlegend = TRUE,
                      xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                      yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
```
</details>

```{r}
# Render the plot
fig_from_paper
```



# 11. Comparison of AI Knowledge Levels Over Time

This analysis compares the distribution of AI knowledge levels between our current data and previous findings, providing insights into how perceptions have evolved over time.

## Overview of Previous Findings

The previous study presented AI knowledge levels as follows:
- **Very Poorly Informed/Uninformed**: 10%
- **Poorly Informed**: 35%
- **Sufficiently Informed**: 36%
- **Well Informed**: 17%
- **Very Well Informed**: 2%

These results indicate a larger proportion of students either poorly informed or only sufficiently informed about AI.

## Current Study Results

Our study shows a different distribution:
- **Very Poorly Informed/Uninformed**: 7.7%
- **Poorly Informed**: 5.5%
- **Sufficiently Informed**: 37.4%
- **Well Informed**: 16.5%
- **Very Well Informed**: 33.0%

These results suggest a significant increase in the proportion of students who are very well informed about AI, coupled with a decrease in those who are very poorly informed or uninformed.

## Visual Comparison

The following visual comparison illustrates these changes:

<details>
<summary>Click here to view the code used for generating the comparison plot</summary>
```{r}
selected_columns2 <- c("Q1.AI_knowledge", "Q3.2.Job_replacement", "Q3.3.Problem_solving", "Q3.4.AI_rulling_society", "Q12.Gender")

data3 <- data[,selected_columns2]


data3 <- data3 %>%
  mutate(Q1.AI_knowledge = case_when(
    Q1.AI_knowledge %in% c(9, 10) ~ "very well informed",
    Q1.AI_knowledge %in% c(7, 8) ~ "well informed",
    Q1.AI_knowledge %in% c(5, 6) ~ "sufficiently informed",
    Q1.AI_knowledge %in% c(3, 4) ~ "poorly informed",
    Q1.AI_knowledge %in% c(1, 2) ~ "very poorly informed/uninformed"
  ),
  Q12.Gender = case_when(
    Q12.Gender == 1 ~ 'F',
    Q12.Gender == 2 ~ 'M',
    TRUE ~ as.character(Q12.Gender)
  ),
  Q3.2.Job_replacement = case_when(
    Q3.2.Job_replacement %in% c(1) ~ "Strongly Disagree",
    Q3.2.Job_replacement %in% c(2) ~ "Partially Disagree",
    Q3.2.Job_replacement %in% c(3) ~ "Neutral",
    Q3.2.Job_replacement %in% c(4) ~ "Partially Agree",
    Q3.2.Job_replacement %in% c(4,5) ~ "Fully Agree"
  ),
  Q3.3.Problem_solving = case_when(
    Q3.3.Problem_solving %in% c(1) ~ "Strongly Disagree",
    Q3.3.Problem_solving %in% c(2) ~ "Partially Disagree",
    Q3.3.Problem_solving %in% c(3) ~ "Neutral",
    Q3.3.Problem_solving %in% c(4) ~ "Partially Agree",
    Q3.3.Problem_solving %in% c(5) ~ "Fully Agree"
  ),
  Q3.4.AI_rulling_society = case_when(
    Q3.4.AI_rulling_society %in% c(1) ~ "Very low",
    Q3.4.AI_rulling_society %in% c(2) ~ "Low",
    Q3.4.AI_rulling_society %in% c(3) ~ "Neighter low or high",
    Q3.4.AI_rulling_society %in% c(4) ~ "High",
    Q3.4.AI_rulling_society %in% c(5) ~ "Very high"))

data3 <- data3 %>%
  mutate(Q1.AI_knowledge = factor(Q1.AI_knowledge, levels = c("very poorly informed/uninformed", "poorly informed", "sufficiently informed", "well informed", "very well informed"), ordered = FALSE))

ordered_levels <- c("Strongly Disagree", "Partially Disagree", "Neutral", "Partially Agree", "Fully Agree")
ordered_levels2 <- c("Very low", "Low", "Neighter low or high", "High", "Very high")

data3 <- data3 %>%
  mutate(Q3.2.Job_replacement = factor(Q3.2.Job_replacement, levels = ordered_levels, ordered = TRUE))
data3 <- data3 %>%
  mutate(Q3.3.Problem_solving = factor(Q3.3.Problem_solving, levels = ordered_levels, ordered = TRUE))
data3 <- data3 %>%
  mutate(Q3.4.AI_rulling_society = factor(Q3.4.AI_rulling_society, levels = ordered_levels2, ordered = TRUE))


knowledge_summary <- data3 %>%
  count(Q1.AI_knowledge) %>%
  rename(Category = Q1.AI_knowledge, Count = n) %>%
  mutate(Percentage = Count / sum(Count))

Distribution_plot <- ggplot(knowledge_summary, aes(x = "", y = Percentage, fill = Category)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  labs(title = "Distribution of AI Knowledge Levels",
       x = NULL, y = NULL, fill = "AI Knowledge Level") +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_blank(),
        plot.title = element_text(size = 16, face = "bold")) +
  scale_fill_brewer(palette = "RdYlGn") + 
  geom_text(aes(label = percent(Percentage)), position = position_stack(vjust = 0.5))
```
</details>

```{r}
print(Distribution_plot)
```

# 12. Job replecemnt by AI

This analysis explores how attitudes towards AI replacing jobs vary by gender, comparing current study results with historical data. This helps us understand trends in perceptions about AI's impact on employment and how these perceptions may differ between males and females.

## Overview of Previous Study Findings

In the previous study, the distribution of attitudes towards AI replacing jobs by gender was as follows:

- **Males** were more likely to fully agree (28.1%) that AI should replace jobs.
- **Females** showed more caution with a lower percentage (5.1%) fully agreeing to the idea.
- **Neutral** positions were significantly less common among males (6.2%) compared to females (22%).

These figures indicate a gender disparity in how the implications of AI on jobs are perceived, with males generally showing more acceptance of AI job replacement.

## Current Study Results

Our study shows a different pattern:

- **Fully Agree**: Males (28.1%) continue to be more accepting than females (5.1%).
- **Partially Agree**: More females (39%) agree to some extent compared to males (31.2%).
- **Neutral**: A significant proportion of females (22%) remain neutral, much higher than males (6.2%).

The table below details these distributions:
<details>
<summary>Click here to view the code and generated table</summary>
```{r, echo=TRUE,warning=FALSE,message=FALSE,error=FALSE}

# Summarize the data by gender and agreement level
table_data2 <- data3 %>%
  group_by(Q12.Gender, Q3.2.Job_replacement) %>%
  summarise(Count = n()) %>%
  mutate(Percentage = Count / sum(Count) * 100) %>%
  arrange(Q12.Gender, Q3.2.Job_replacement)


# Pivot for a side-by-side comparison
pivot_table2 <- table_data2 %>%
  select(-Count) %>%
  pivot_wider(names_from = Q12.Gender, values_from = Percentage, values_fill = list(Percentage = 0))

pivot_table2 <- pivot_table2 %>%
  mutate(across(where(is.numeric), ~ paste0(round(.x, 1), "%")))
```
</details>

```{r}
# Display the table
pivot_table2 %>% 
  kable( format = "markdown", col.names = c("Replacing the humans and jobs by gender", "Female (%)", "Male (%)")) %>% 
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

### Comparing the two data sets, we see that:

-  Students in 2022 are less willing to fully commit to one side of the argument about AI replacing jobs. There's a general trend towards more neutral and partially agreeing positions, especially among females.
-  The increase in the "Partially Agree" category among females suggests a growing complexity in how job replacement by AI is viewed, possibly reflecting broader societal and economic discussions about AI.



# 13. Comparative Analysis of Perceptions on AI as a Threat to Humanity by Gender

This section examines the differences in perceptions about whether AI could rule society or destroy humanity, comparing our current study results with historical data to see how these views have evolved over time.

### Overview of Previous Study Findings

The previous study showed a notable gender difference in the perception of AI's threat:

- **Females** expressed higher concern with 15.5% perceiving a very high threat and 29.5% a high threat.
- **Males** were less concerned, with only 8.4% perceiving a very high threat and 22.7% a high threat.

These results suggest a higher level of apprehension among females regarding the potential dangers of AI.

### Current Study Results

Our study indicates a shift in these perceptions:

- **Very Low Threat**: More females (30.5%) now perceive a very low threat compared to males (31.2%).
- **High Threat**: Fewer females now see AI as a high threat (8.5%) compared to males (18.8%).

The table below details these distributions:

<details>
<summary>Click here to view the code and generated table</summary>

```{r, echo=TRUE,warning=FALSE,message=FALSE,error=FALSE}
library(knitr)
library(kableExtra)

# Summarize the data by gender and perception level
table_data3 <- data3 %>%
  group_by(Q12.Gender, Q3.4.AI_rulling_society) %>%
  summarise(Count = n()) %>%
  mutate(Percentage = Count / sum(Count) * 100) %>%
  arrange(Q12.Gender, Q3.4.AI_rulling_society)

# Pivot for a side-by-side comparison
pivot_table3 <- table_data3 %>%
  select(-Count) %>%
  pivot_wider(names_from = Q12.Gender, values_from = Percentage, values_fill = list(Percentage = 0))

pivot_table3 <- pivot_table3 %>%
  mutate(across(where(is.numeric), ~ paste0(round(.x, 1), "%")))


```
</details> 

### Discussion
- **Decrease in Perceived Threat**: The data from 2022 indicate a general decrease in the perception of AI as a severe threat to humanity, especially among females.
- **Gender Differences**: While both genders show a reduction in concern, males still report higher levels of concern compared to females, reversing the trend seen in previous data.

This comparative analysis highlights how societal attitudes towards AI and its potential risks are changing, possibly reflecting increased familiarity and integration of AI technologies in daily life.

```{r}
# Display the table
pivot_table3 %>% 
  kable(format = "markdown", col.names = c("AI will rule society/destroy humanity", "Female (%)", "Male (%)")) %>% 
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```
# 14. Comparative Analysis of Overall Attitudes Towards AI and Its Consequences

This section examines how attitudes towards the positive impacts of AI on society differ by gender and compares current study results with those from historical data.

## Overview of Previous Study Findings

In the previous study, attitudes towards AI's potential to positively impact society showed:

- **Males** were significantly more optimistic with 66.8% believing in a positive influence of AI.
- **Females** were less optimistic with only 51.5% sharing that belief.

This data suggests that males tend to view AI development more favorably than females.

## Current Study Results

Our current study presents a shift in these attitudes:

- **Highly Positive** views are less common in the current data, with only 1.7% of females and 12.5% of males seeing AI's impact as 'Very high'.
- **Neutral to Negative** views are more prevalent, especially among females, with 30.5% viewing the impact as 'Very low'.

The table below details these distributions:

<details>
<summary>Click here to view the code and generated table</summary>


```{r, echo=TRUE,warning=FALSE,message=FALSE,error=FALSE}
# Summarize the data by gender and views on AI's problem-solving capabilities
table_data4 <- data3 %>%
  group_by(Q12.Gender, Q3.3.Problem_solving) %>%
  summarise(Count = n()) %>%
  mutate(Percentage = Count / sum(Count) * 100) %>%
  arrange(Q12.Gender, Q3.3.Problem_solving)
# Pivot for a side-by-side comparison
pivot_table4 <- table_data4 %>%
  select(-Count) %>%
  pivot_wider(names_from = Q12.Gender, values_from = Percentage, values_fill = list(Percentage = 0))
# Pivot for a side-by-side comparison
pivot_table4 <- pivot_table4 %>%
  mutate(across(where(is.numeric), ~ paste0(round(.x, 1), "%")))


```
</details> 

```{r}

# Display the table
pivot_table3 %>% 
  kable( format = "markdown", col.names = c("AI helps to solve many problems in society", "Female (%)", "Male (%)")) %>% 
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```
### Discussion
- **Decrease in Optimism:** Students in 2022 are less optimistic about AI having a positive influence on society compared to previous years. This trend is more pronounced among females.
- **Gender Differences:** While both genders show less optimism, the decline is steeper among females, suggesting that experiences or perceptions of AI's societal impact may differ significantly by gender.



# 15 Tests

Hypothesis 2 (H2). Male students have a different level of information regarding AI use than female students.

## 15.1 Levene's test for Equality of variances
```{r}
data$Q12.Gender <- as.factor(data$Q12.Gender)
levene <- leveneTest(Q1.AI_knowledge ~ Q12.Gender, data = data)
levene_f <- levene$`F value`[1]
levene_p <- levene$`Pr(>F)`[1]
```

## 15.2 t-Test for equality of means
```{r}
t_test_equal_var <- t.test(Q1.AI_knowledge ~ Q12.Gender, data = data, var.equal = TRUE)
t_equal <- t_test_equal_var$statistic
df_equal <- t_test_equal_var$parameter
p_equal <- t_test_equal_var$p.value
mean_diff_equal <- t_test_equal_var$estimate[1] - t_test_equal_var$estimate[2]
```

```{r}
t_test_unequal_var <- t.test(Q1.AI_knowledge ~ Q12.Gender, data = data)
t_unequal <- t_test_unequal_var$statistic
df_unequal <- t_test_unequal_var$parameter
p_unequal <- t_test_unequal_var$p.value
mean_diff_unequal <- t_test_unequal_var$estimate[1] - t_test_unequal_var$estimate[2]
```

```{r}
original_table <- data.frame(
  ` ` = c("Equal variances assumed", "Equal variances not assumed"),
  F = c(5.7, ""),
  Sig = c(0.017, ""),
  t = c(9.269, 9.254),
  df = c(925, 886.056),
  `Sig (2-Tailed)` = c(0.000, 0.000),
  `Mean Difference` = c(0.553, 0.553)
)

# Display the table using kable
original_table %>% kable(col.names = c("", "F", "Sig", "t", "df", "Sig (2-Tailed)", "Mean Difference"), caption = "Independent samples test for H2.") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

```{r}
summary_table <- data.frame(
  ` ` = c("Equal variances assumed", "Equal variances not assumed"),
  F = c(levene_f,''),
  Sig = c(levene_p,''),
  t = c(t_equal, t_unequal),
  df = c(df_equal, df_unequal),
  "Sig (2-Tailed)" = c(p_equal, p_unequal),
  "Mean difference" = c(mean_diff_equal, mean_diff_unequal)
)

# Print summary table
summary_table %>% kable( format = "markdown") %>% 
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```


Tests indicate that:
- In both cases difference in variance is significant
- In original data difference in AI knowledge between male and male students is significant with difference of .553 while in our data difference is equal only to -.0874
